---
title: "KaizoIronmon"
author: "Celine Mangahas"
date: "2024-04-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Pokemon has been one of the most popular games for the last 28 years. However, some players have agreed the recent games have dropped in difficulty; they needed a new challenge, a better twist on their favorite game series. Iateyourpie created a Pokemon Randomizer that randomizes Pokemon spawns, stats, items, even abilities to push players to the limit: The IronMON Challenge. One of the most famous versions of the IronMON Challenge is Kaizo IronMON—with 38 rules restricting the player on what items, Pokemon, moves and abilities one can use against harder, smarter AI. It's no wonder the challenge has many losses and very, very few wins.

There are various analyses from the moderators and streamers that have taken on the challenge, however, using the R programming language could create a new perspective for all to see. Graphs such as the violin and dodge plots could be useful to see which are the most popular/used types for the challenge, and possibly using the package "tidyverse" for data visualization. I would also like to evaluate the relationship between abilities and Pokemon types to see how valuable typing and abilities are. I could also look at how certain stats could play into a gamer winning their challenge. I would have to webscrape the data borrowed from the IronMON community with permission.  
```{r ironmon1, echo=FALSE}
ironmon <- read.csv("FRLG_kaizo_ironmon.csv")
View(ironmon)
```

## Graph/Visual Analysis of Kaizo Ironmon

Win status by Pokemon:

```{r ironmon2, echo=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)


# Bar plot of Win status by Pokemon
ggplot(ironmon, aes(x = Pokemon, fill = Win)) +
  geom_bar() +
  labs(title = "Win Status by Pokemon", x = "Pokemon", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Facet grid of Type1 vs Special Attack color-organized by Win Status
```{r ironmon3, echo=FALSE}
# Facet grid for Type1 vs Special Attack with Win Status
ggplot(ironmon, aes(x = Type1, y = SpecialAttack, fill = Win)) +
  geom_boxplot() +
  labs(title = "Facet Grid: Type1 vs Special Attack with Win Status",
       x = "Type1",
       y = "Special Attack",
       fill = "Win") +
  facet_wrap(~ Win) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Box plot of Win statuses:
```{r ironmon5, echo=FALSE}
# Box plot of BST by Win status
ggplot(ironmon, aes(x = Win, y = BST, fill = Win)) +
  geom_boxplot() +
  labs(title = "BST by Win Status", x = "Win", y = "BST") +
  theme_minimal()
```
Scatter plot of Special attack vs. Defense. There is a smoothed line on the graph to represent a possible slope to show the relationship between special attack and defense, however, due to the random number generation (RNG for short in gamer terminology), there is no general relationship detected:
```{r ironmon6, echo=FALSE}
# Scatter plot with smoothed line
ggplot(ironmon, aes(x = SpecialAttack, y = Defense)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Scatter Plot: Special Attack vs Defense",
       x = "Special Attack",
       y = "Defense") +
  theme_minimal()
```
Filtered Data of Pokemon with BST under 500
```{r ironmon7, echo=FALSE}
# Filter data for Pokemon with BST under 500
pokemon_bst_under_500 <- ironmon %>% 
  filter(BST < 500)

# Bar plot of Wins for Pokemon with BST under 500
ggplot(pokemon_bst_under_500, aes(x = factor(Win), fill = factor(Win))) +
  geom_bar() +
  labs(title = "Wins for Pokemon with BST under 500",
       x = "Win",
       y = "Count") +
  scale_fill_manual(values = c("red", "blue")) +  # Colors for wins
  theme_minimal()
```
Filtered Data of Pokemon with BST under 450
```{r ironmon8, echo=FALSE}
# Assuming ironmon is your data frame
pokemon_bst_under_450 <- ironmon %>% 
  filter(BST < 450)

# View the filtered data
head(pokemon_bst_under_450)

# Bar plot of BST under 450 vs. Wins
ggplot(pokemon_bst_under_450, aes(x = Pokemon, fill = Win)) +
  geom_bar() +
  labs(title = "Wins vs. BST under 450 for All Types",
       x = "Pokemon",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Bar chart of which types win the most. Water and normal types seem to win the most from my analyses.
```{r ironmon9, echo=FALSE}
# Bar plot of Wins by Type1
ggplot(ironmon, aes(x = Type1, fill = Win)) +
  geom_bar() +
  labs(title = "Wins by Type1",
       x = "Type1",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Stacked bar plot of Wins by Type1 and Type2
ggplot(ironmon, aes(x = Type1, fill = Type2)) +
  geom_bar(position = "stack") +
  labs(title = "Wins by Type1 and Type2",
       x = "Type1",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = guide_legend(title = "Type2"))
```
Stacked Bar chart of which types win the most. Water and normal types seem to win the most from my analyses.
```{r ironmon10, echo=FALSE}
# Stacked bar plot of Wins by Type1 and Type2
ggplot(ironmon, aes(x = Type1, fill = Type2)) +
  geom_bar(position = "stack") +
  labs(title = "Wins by Type1 and Type2",
       x = "Type1",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = guide_legend(title = "Type2"))
library(dplyr)

# Calculate total wins and losses by Type1 and Type2
wins_losses <- ironmon %>%
  group_by(Type1, Type2) %>%
  summarize(TotalWins = sum(Win == "TRUE"),
            TotalLosses = sum(Win == "FALSE"))

# View the wins_losses table
print(wins_losses)
```
## Numerical Analysis of Ironmon
To determine whether or not the data fits a normal distribution (data is symmetric to the mean, showing data near the mean to be more frequent closer than farther), I used the Shapiro-Wilk test to find the p-value of the following values: IV, HP, Speed, Special Defense, Special Attack, Attack, Base Stat, Level, and Defense.

Let H0, our null hypothesis, be each value follow a Normal distribution
Let Ha, the alternative hypothesis, - each of the values does not follow a Normal distribution
```{r ironmon11, echo=FALSE}
# Shapiro-Wilk test for IV
shapiro_test_IV <- shapiro.test(ironmon$IV)
shapiro_test_IV

# Shapiro-Wilk test for Speed
shapiro_test_speed <- shapiro.test(ironmon$Speed)
shapiro_test_speed

# Shapiro-Wilk test for Attack
shapiro_test_atk <- shapiro.test(ironmon$Attack)
shapiro_test_atk

# Shapiro-Wilk test for Defense
shapiro_test_def <- shapiro.test(ironmon$Defense)
shapiro_test_def

# Shapiro-Wilk test for HP
shapiro_test_HP <- shapiro.test(ironmon$HP)
shapiro_test_HP

# Shapiro-Wilk test for Special Defense
shapiro_test_spD <- shapiro.test(ironmon$SpecialDefense)
shapiro_test_spD

# Shapiro-Wilk test for Special Attack
shapiro_test_spA <- shapiro.test(ironmon$SpecialAttack)
shapiro_test_spA

# Shapiro-Wilk test for Special Attack
shapiro_test_lvl <- shapiro.test(ironmon$PokemonLvl)
shapiro_test_lvl

# Shapiro-Wilk test for Base Stat
shapiro_test_bst <- shapiro.test(ironmon$BST)
shapiro_test_bst
```
Whenever the p-value is above 0.05, we reject H0 and accept the HA as our hypothesis. IV and Speed are rejected due to the randomness of the values from our sample, but the other values seem to follow a normal distribution.

## Wins and Losses Analyses
Here is a calulation of wins and losses by types.
```{r ironmon12, echo=FALSE}
# Calculate total wins and losses by Type1 and Type2
wins_losses <- ironmon %>%
  group_by(Type1, Type2) %>%
  summarize(TotalWins = sum(Win == "TRUE"),
            TotalLosses = sum(Win == "FALSE"))

# View the wins_losses table
print(wins_losses)

# Type1 and Type2 with most wins
most_wins <- wins_losses %>%
  filter(TotalWins == max(TotalWins))

# Type1 and Type2 with most losses
most_losses <- wins_losses %>%
  filter(TotalLosses == max(TotalLosses))

# Print the results
cat("Type1 and Type2 with the most wins:\n")
print(most_wins)

cat("\nType1 and Type2 with the most losses:\n")
print(most_losses)
```

```{r ironmon13, echo=FALSE}

# Calculate counts of wins and losses
win_counts <- sum(ironmon$Win == TRUE)
loss_counts <- sum(ironmon$Win == FALSE)

# Create a data frame for the pie chart
pie_data <- data.frame(Win = c("Wins", "Losses"),
                       Count = c(win_counts, loss_counts))

# Create the pie chart
ggplot(pie_data, aes(x = "", y = Count, fill = Win)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Pie Chart of Win Status",
       fill = "Win") +
  theme_void() +
  theme(legend.position = "right") +
  geom_text(aes(label = paste0(round(Count/sum(Count) * 100), "%")),
            position = position_stack(vjust = 0.5))

# Filter data for Pokemon that won
winning_pokemon <- ironmon %>%
  filter(Win == TRUE)

# Count occurrences of each Type1 for winners
win_type1_counts <- count(winning_pokemon, Type1)

# Calculate total wins by Type1
total_wins_type1 <- ironmon %>%
  filter(Win == TRUE) %>%
  count(Type1, sort = TRUE)

# View the total wins by Type1
print(total_wins_type1)
```